---
title: "STAT425-HW3"
author: "Aidana Karipbayeva"
date: "10/13/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
```{r}
library(faraway)
data(teengamb)
model1 <- lm(gamble ~ ., data = teengamb)
summary(model1)

```

### Q1. a.
```{r}
par(mar=c(1,1,1,1))
plot(fitted(model1), resid(model1))

```
From the above plot, we can see that the constant variance assumption for the errors is violated, as the points form a megaphone shape which can suggest us to modify the chosen model.

### Q1. b.
```{r}
qqnorm(resid(model1))
qqline(resid(model1))
```
It seems like the errors are normally distributed as most of the points lie on the line.


### Q1. c.
```{r}
plot(hatvalues(model1))
which(hatvalues(model1) >= 10/47)
```
From the above plot, we can see that there are 4 points with high leverage. Testing points with leverage greater than 2p/n confirms that 4 observation points (31 33 35 42) have high leverage


### Q1. d.

```{r}
plot(rstudent(model1))

```
From the rstudentized plot, we can see that there is one outlier as its residual is much larger than other observations' residuals.

### Q1. e.
```{r}
plot(cooks.distance(model1))
```
From the plot, we can see that there is no observation with Cook's distance bigger than 1. Thus, we can conclude that there are no influential points.

## Question 2

### Q2. a.1
```{r}
set.seed(1)
x <- rnorm(n = 100)

```

### Q2. b.
```{r}
plot(sort(x), sort(x), xlab = "Theoretical Quantiles", ylab = "Simple Quantiles", main= "My Q-Q Plot")

```

## Question 3
```{r}
parent <- 15:21
offspring <- c(15.89, 16.16, 16.12, 16.39, 16.36, 17.06, 17.25)
sd = c(1.764, 1.595, 1.655, 2.036, 1.895, 1.937, 1.987)
Gru <- data.frame(parent, offspring, sd)
```

### Q3. a.
```{r}
X=parent
Y=offspring
plot(X, Y)
```
### Q3. b.
```{r}
ols <- lm(Y~X)
summary(ols)
```
### Q3. c.
```{r}
wls <- lm(Y ~ X, weights=sd^-2)
summary(wls)
```

### Q3. d.
```{r}
plot(X, Y, col = "blue",main = "Fitted lines from OLS and WLS", xlab = "Parent", ylab = "Offspring")
abline(ols, col = "red")
abline(wls, col = "green")
```
### Q3. e.
```{r}
t_test <- (wls$coefficients[2]-0.5)/coef(summary(wls))[, "Std. Error"][2]
p_value <- pt(q = t_test, df = 5)
```
Test statistic is equal to `r t_test`, its distibution under H0 follows t-distribution and p-value is equal to `r p_value`. It is less than 0.05(my chosen alpha), so we will reject the null hypothesis.


## Question 4
```{r}
library(faraway)
data(strongx)
```
### Q4. a. 
```{r}
str_model = lm(crossx ~ energy, data = strongx)
corr = cor(str_model$res[-1], str_model$res[-10])
S = diag(10)
Sig.mat <- round(corr ^ abs(row(S) - col(S)),2)
Sig.mat
Sig.inv <- round(solve(Sig.mat),2)
Sig.inv

```

### Q4. b.
```{r}
X = model.matrix(str_model)
beta_hat<-solve(t(X)%*%Sig.inv%*%X)%*%t(X)%*%Sig.inv%*%strongx$crossx
beta_hat

```

## Question 5
```{r}
golf <- read.table('lpga2009.csv', header = TRUE, sep = ",")
```

## Q5. a.
```{r}
model3 <- lm(prize ~ percentile, data = golf)
summary(model3)
quartz.save();par(mfrow=c(2,2));plot(model3)
```

## Q5. b.
It seems like residuals are normally distributed from Q-Q plot, as points lie on the line.


## Q5. c.
```{r}
outliers <- which(rstandard(model3) > 2.5)
plot(rstudent(model3))
outliers

```
From the plot, we can see that there is one big outlier and a few smaller outliers. The observation numbers of these outliers are `r outliers`


## Q5. d.
From the Residuals vs Leverage plot, we can see that there are no influential points.


## Q5. e.
```{r}
library(lmtest)
bptest(model3)
```
We reject the null hypothesis that there is a homoscedasticity because p-value is significant (less than 0.05).
